### 2017-06-12 Draft 2/Section 1 - Lambda Calculus.tex

\documentclass[Master.tex]{subfiles}

\begin{document}

The simplest, and in my opinion the most beautiful definition of effective calculability was introduced by the American mathematician Alonzo Church in the 1930s, as a purely mathematical and abstract definition of computability. It deals with a set of objects - for our purposes denoted by single symbolic characters - which define what school mathematics or common computer science would call `functions'. These functions take as their arguments (other objects which the functions act upon) functions alone, and can also output only functions. Take, for example the function $f$:
\begin{equation*}
f = ab
\end{equation*}
This particular function $f$ `applies' the function $a$ to the function $b$. Note that while the syntax $a(b)$ would normally be used, it greatly simplifies expression if the brackets are omitted. Church defines these functions through the method of `abstraction' - to illustrate:
\begin{equation*}
\lambda u.tu
\end{equation*}
This function takes one single argument, which it names the abstract name $u$. It then applies the function $t$ to $u$ and outputs the result. For example, if this was to be applied to the function $a$:
\begin{equation*}
(\lambda u.tu) a = ta	
\end{equation*}

Functions can take multiple arguments through the use of nesting:
\begin{gather*}
\begin{aligned}
\lambda a.(\lambda b.ba) &=  \lambda ab.ba \\
[\lambda a.(\lambda b.ba)]uv &= (\lambda b.bu)v \\ &= vu
\end{aligned}
\end{gather*}

\subsection{Arithmetical definitions}

The most basic entities that can be defined in this medium are the natural numbers:
\cite{church1941lambda}
\begin{equation*}
\begin{aligned}
\bm{\mathrm{0}} &= \lambda sz.z,
  & % seperation
\bm{\mathrm{1}} &= \lambda sz.sz,
  \\
\bm{\mathrm{2}} &= \lambda sz.s(sz),
  &
\bm{\mathrm{3}} &= \lambda sz.s(s(sz)),
  &
etc.
\end{aligned}
\end{equation*}
Here it can be seen that for each number \textit{n} the argument $s$ is applied to the argument $z$, \textit{n} times - the $s$ functions representing \textit{n} successors to zero ($z$).

From this, it is not difficult to define an operation that increments a given number by 1. Such a function takes an argument $w$, representing the number to be incremented, and returns a number which is equal to $w$ with the $s$ argument (here represented by $y$) applied to it once.
\cite{rojas2015lambdatutorial}
\begin{equation*}
\begin{aligned}
\bm{\mathrm{S}} &= \lambda w.\lambda yx.y(wyx)\\
&= \lambda wyx.y(wyx)
\end{aligned}
\end{equation*}
Let us test our new successor function by applying it to \textbf{2}:
\begin{gather*}
\begin{aligned}
\bm{\mathrm{S2}} &= (\lambda wyx.y(wyx)) (\lambda sz.s(s(z)))\\
&= \lambda yx.y([\lambda sz.(s(s(z)))]yx)\\
&= \lambda yx.y(y(y(z)))\\
&= \lambda sz.s(s(s(z)))\\
&= \bm{\mathrm{3}}
\end{aligned}
\end{gather*}

Observing that each number applies the function $s$ to $z$ \textit{n} times, and that addition is simply incrementing \textit{a} \textit{b} times, addition is as simple as:
\cite{rojas2015lambdatutorial}
\begin{equation*}
a\bm{\mathrm{S}}b	
\end{equation*}
This can be tested by summing \textbf{2} and \textbf{3}:
\begin{gather*}
\begin{aligned}
\bm{\mathrm{2S3}} &= (\lambda sz.s(s(z))) \bm{\mathrm{S}} \bm{\textrm{3}}\\
&= \bm{\mathrm{S}} (\bm{\mathrm{S}} (\bm{\mathrm{3}}))\\
&= \bm{\mathrm{S}} (\bm{\mathrm{4}})\\
&= \bm{\mathrm{5}}
\end{aligned}
\end{gather*}

\subsection{Boolean Logic}

Church defines functions to operate boolean logic, fundamentally defining True and False to be:
\cite{rojas2015lambdatutorial}

\begin{equation*}
\begin{aligned}
\bm{\mathrm{T}} &= \lambda ab.a
  & % seperation
\bm{\mathrm{F}} &= \lambda ab.b
\end{aligned}
\end{equation*}
These functions each take two arguments - the \textbf{T} function returns the first, while the \textbf{F} function returns the second.

These concepts are the base of conditional branching in computation, and in $\lambda$-Calculus operate like \textit{if} statements found in most programming languages. For such operations, the concept of a `pair' exists:
\cite{church1941lambda}
\begin{equation*}
\lambda x.xab
\end{equation*}
where $a$ and $b$ are functions. For a choice of an element from the pair, the argument of either \textbf{T} or \textbf{F} is provided:
\begin{gather*}
\begin{aligned}
(\lambda x.xab)\bm{\mathrm{T}} &= a\\
(\lambda x.xab)\bm{\mathrm{F}} &= b
\end{aligned}
\end{gather*}

Functions representing boolean operations can also be defined:
\cite{rojas2015lambdatutorial}
\begin{gather*}
\begin{aligned}
\vee &= \lambda xy.x\bm{\mathrm{T}}y\\
\wedge &= \lambda xy.xy\bm{\mathrm{F}}\\
\lnot &= \lambda x.x\bm{\mathrm{FT}} \\
\veebar &= \lambda xy.x(\lnot y)y
\end{aligned}
\end{gather*}
The derivation of such operations is not difficult to understand if one considers the truth tables for each operation, remembering that T selects the first and F selects the second argument - for example, looking at the OR function $\vee$:

\begin{gather*}
\begin{aligned}
\vee \bm{\mathrm{TF}} &= \bm{\mathrm{T}}[\bm{\mathrm{T}}][\bm{\mathrm{F}}] \\
&= \bm{\mathrm{T}} \\
\vee \bm{\mathrm{FF}} &= \bm{\mathrm{F}}[\bm{\mathrm{T}}][\bm{\mathrm{F}}] \\
&= \bm{\mathrm{F}} \\
\vee \bm{\mathrm{FT}} &= \bm{\mathrm{F}}[\bm{\mathrm{T}}][\bm{\mathrm{T}}] \\
&= \bm{\mathrm{T}} \\
\end{aligned}
\end{gather*}

If the first argument is \textbf{T} the function will always return \textbf{T}, as it is the first argument. If the first argument is \textbf{F} on the other hand, the value of the second argument will be returned (the $\vee$ function is dependant on at least one argument being \textbf{T} - if this is not the first argument, the outcome is simply whether the second argument is \textbf{T} or not). The remaining functions work in a similar way.

\subsection{Basic Comparison operators}

With the fundamental boolean system of our computational system in place, we will next construct comparison operators. The first and simplest of these will test if a number is equal to zero:
\cite{rojas2015lambdatutorial}
\begin{equation*}
\bm{\mathrm{Z}} = \lambda x.x\bm{\mathrm{F\lnot F}}
\end{equation*}
To understand this somewhat less than trivial concept, note that \textbf{F} when given only a single argument yields the identity function $\lambda x.x$, and that \textbf{0} applied to any two arguments yields only the second:
\cite{rojas2015lambdatutorial}
\begin{equation*}
\bm{0}ab = (\lambda sz.z)ab = b
\end{equation*}
From this, it easier to see how the function works. Firstly with \textbf{0}:
\cite{rojas2015lambdatutorial}
\begin{gather*}
\begin{aligned}
\bm{\mathrm{Z0}} &= (\lambda x.x\bm{\mathrm{F\lnot F}})\bm{0}\\
&= \bm{\mathrm{0F\lnot F}}\\
&= \bm{\mathrm{\lnot F}}\\
&= \bm{\mathrm{T}}
\end{aligned}
\end{gather*}
Secondly with \textbf{2}:
\begin{gather*}
\begin{aligned}
\bm{\mathrm{Z2}} &= (\lambda x.x\bm{\mathrm{F\lnot F}})\bm{2}\\
&= \bm{\mathrm{2F\lnot F}}\\
&= (\bm{\mathrm{F}}(\bm{\mathrm{F}}\lnot)) \bm{\mathrm{F}}\\
&= (\bm{\mathrm{F}}(\lambda x.x))\bm{\mathrm{F}}\\
&= (\lambda x.x)\bm{\mathrm{F}}\\
&= \bm{\mathrm{F}} 
\end{aligned}
\end{gather*}

However, in order to develop other conditional tests we will need a `predecessor' function - a more complex concept in $\lambda$-Calculus.
\subsection{Subtraction}

The elusive predecessor function was created by Stephen Kleene, a student of Church's, and was something of a landmark in the development of $\lambda$-Calculus.

Firstly, the function $\bm{\psi}$ transforms the pair (n, n-1) (the argument $p$) into the pair (n+1, n) (the new pair's elements are an increment and a copy of the first value respectively):
\cite{rojas2015lambdatutorial}
\begin{equation*}
\bm{\psi} = \lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})
\end{equation*}
To find the predecessor to $n$ therefore, we apply this function $n$ times to the pair (0, 0) and the pair (n, n-1) is formed, after which the second element can easily be selected.
\cite{rojas2015lambdatutorial}
\begin{equation*}
\bm{\mathrm{P}} = \lambda n.[n \bm{\psi} (\lambda x.\bm{\mathrm{00}})] \bm{\mathrm{F}} 
\end{equation*}
\begin{equation*}
\begin{aligned}
\bm{\mathrm{P3}} &= (\lambda n.n [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})] [\lambda x.\bm{\mathrm{00}}] \bm{\mathrm{F}}) [\lambda sz.s(s(s(z)))]\\
&= ([\lambda sz.s(s(s(z)))] [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})] [\lambda x.\bm{\mathrm{00}}]) \bm{\mathrm{F}}\\
&= ([\lambda sz.s(s(s(z)))] [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})] [\lambda x.\bm{\mathrm{00}}]) \bm{\mathrm{F}}\\
&= [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})]([\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})]([\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})][\lambda x.\bm{\mathrm{00}}]))\bm{\mathrm{F}}\\
&= [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})]([\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})](\lambda c.c(\bm{\mathrm{S}}[\lambda x.\bm{\mathrm{00}}]\bm{\mathrm{T}})([\lambda x.\bm{\mathrm{00}}]\bm{\mathrm{T}})))\bm{\mathrm{F}}\\
&= [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})]([\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})](\lambda c.c\bm{\mathrm{10}}))\bm{\mathrm{F}}\\
&= [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})][\lambda c.c\bm{\mathrm{21}}]\bm{\mathrm{F}}\\
&= [\lambda c.c\bm{\mathrm{32}} ])\bm{\mathrm{F}}\\
&= \bm{\mathrm{2}}
\end{aligned}
\end{equation*}

Of course now the subtraction function can be found (using \textbf{D} for difference):
\begin{equation*}
\bm{\mathrm{D}} = \lambda ab.b\bm{\mathrm{P}}a
\end{equation*}
Bearing in mind the predecessor of \textbf{0} is \textbf{0}, this will produce \textbf{0} if $b\geq a$.
\subsection{Further Conditional Logic}

With this useful quirk of the \textbf{D} function, we can produce comparative logical operators:
\begin{equation*}
\begin{aligned}
\bm{\leq} &= \lambda ab.\bm{\mathrm{Z}}(\bm{\mathrm{D}}ab)
  & % seperation
\bm{\geq} &= \lambda ab.\bm{\mathrm{Z}}(\bm{\mathrm{D}}ba)
  \\
\bm{<} &= \lambda ab.\bm{\mathrm{Z}}(\bm{\mathrm{D}}a[\bm{\mathrm{S}}b])
  &
\bm{>} &= \lambda ab.\bm{\mathrm{Z}}(\bm{\mathrm{D}}[\bm{\mathrm{S}}b]a)
\end{aligned}
\end{equation*}
and through combination of these, we can create the equivalence and inequivalence operators (using $\bm{\doteq}$ for clarity from $=$):
\begin{equation*}
\begin{aligned}
\bm{\doteq} &= \lambda ab.\wedge(\bm{\leq}ab)(\bm{\geq}ab)\\
\bm{\neq} &= \lambda ab.\vee(\bm{<}ab)(\bm{>}ab)
\end{aligned}
\end{equation*}
\subsection{Recursion}

Continuity is one of the most significant aspects of computation, the power of this being the ability to transform a finite configuration into a potentially infinite result. This is done in $\lambda$-Calculus through the use of one of the most remarkable functions, the \textbf{Y} combinator:
\cite{rojas2015lambdatutorial}
\begin{equation*}
\bm{\mathrm{Y}} = \lambda y.(\lambda x.y[xx])\ (\lambda x.y[xx])
\end{equation*}
\begin{equation*}
\begin{aligned}
\bm{\mathrm{Y}}a &= (\lambda x.a[xx])\ (\lambda x.a[xx])\\
&= a\ [(\lambda x.a[xx])\ (\lambda x.a[xx])]\\
&= a\ [\bm{\mathrm{Y}}a]
\end{aligned}
\end{equation*}
This function is remarkable because it is a `fixpoint operator' - it is able to easily generate a `fixpoint' of any given function (a value $x$ is a fixpoint of the function $f$ if and only if $fx = x$) by applying the function as the first parameter of the \textbf{Y} combinator.

Therefore, the \textbf{Y} combinator applied to any function $a$ applies the function $a$ to the function \textbf{Y} applied to $a$ - if substitution continues, an infinite series of nested $a$ functions is created; an infinite result is produced from a finite configuration. Of course we can create finites as well by halting the function - an example of this would be the factorial function, where \textbf{!}a = \textbf{M}a(\textbf{!}[\textbf{P}a]) and \textbf{!0} = \textbf{1}. Firstly, a recursive definition:
\begin{equation*}
\begin{aligned}
\bm{\mathrm{\xi}} &= \lambda fa.\bm{\mathrm{Z}}a\bm{\mathrm{1}}[\bm{\mathrm{M}}a(f[\bm{\mathrm{P}}a])]\\
\bm{!} &= \bm{\xi}!
\end{aligned}
\end{equation*}
i.e. the factorial function is the $\bm{\xi}$ function with itself as the argument $f$. Unfortunately such a definition is illegal under $\lambda$-Calculus as all functions are strictly anonymous - this is where the \textbf{Y} combinator comes into play. To find the factorial of \textbf{3}:
\begin{equation*}
\begin{aligned}
\bm{\mathrm{Y\xi 3}} &= \bm{\mathrm{\xi}}(\bm{\mathrm{Y\xi}})\bm{\mathrm{3}}\\
&= \bm{\mathrm{Z3}}\ [\bm{\mathrm{1}}]\ [\bm{\mathrm{M3}}([\bm{\mathrm{Y\xi}}][\bm{\mathrm{P3}}])]\\
&= \bm{\mathrm{M3}}(\bm{\mathrm{Y\xi}}\bm{2})\\
&= \bm{\mathrm{M3}}(\bm{\mathrm{M2}}(\bm{\mathrm{M11}}))\\
&= \bm{\mathrm{6}}
\end{aligned}
\end{equation*}
The \textbf{Y} combinator allows for \textbf{Y}$\bm{\xi}$ to be applied as the first argument to the $\bm{\xi}$ function - as a result, the procedure also processes \textbf{2} and \textbf{1}, before stopping at \textbf{0} (which produces simply \textbf{1}. To simplify, the factorial function:
\begin{equation*}
\bm{!} = \lambda x.\bm{\mathrm{Y\xi}}x
\end{equation*}
\subsection{Conclusion}

Our brief stroll through the key aspects of $\lambda$-Calculus is complete. We have shown that many basic necessary concepts in a common programming language can be defined through the use of two simple axioms - that of the concept of functions, and the operation of substitution. $\lambda$-Calculus was described by Church as an `effectively calculable' system, in that it could perform any calculable calculation. In more common terms, it is Turing-complete - we will look at what this is and why this is the case later on. However, it is clear to see that the Calculus is able to simulate many functions of the computer - data storage with pairs (pairs of pairs can be used for storing more than two items of data as shown in \cref{sec:LambdaUnivTape}), conditionals and booleans, integers and arithmetic and, finally, the ability to be continuous (circular) from a finite configuration. $\lambda$-Calculus is our first attempt to clarify the attributes of a computer, and does so both simply and aesthetically.

\end{document}